---
title: "Machine Learning Assignment"
author: "Max"
date: "3/25/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation

We load the caret package and the data.

```{r }
library(caret)

set.seed(789)

training_data <- read.csv("pml-training.csv", na.strings=c("NA",""), strip.white=TRUE)

dim(training_data)
```

We need to clean the data by removing columns which are NA. We also remove timestap variables as we do not want to use these to predict.

```{r }
num_of_NA <- apply(training_data, 2, function(x) { sum(is.na(x)) })
training_data <- training_data[, which(num_of_NA == 0)]
training_data <- training_data[, -(1:7)]

dim(training_data)
```

We will partition the data into training and test sets.

```{r }
training_index <- createDataPartition(y=training_data$classe, p=3/4, list=FALSE)
training <- training_data[training_index,]
testing <- training_data[-training_index,]
```


## Training a Random Forest Model

We train a random forest model on the training set.

```{r }
control <- trainControl(allowParallel=TRUE, method="cv", number=4)
model <- train(classe ~ ., data=training, model="rf", trControl=control)
prediction <- predict(model, newdata=testing)
```

We check the predictions against the testing dataset.

```{r }
sum(prediction == testing$classe) / length(prediction)
```

```{r }
confusionMatrix(testing$classe, prediction)$table
```

Our model is 99.2% accurate when apllied to the testing dataset.  The confusion matrix looks good as well. 

We show the most important in our model?

```{r }
varImp(model)
```

## Testing the Model

We now use our model to predict the 20 test cases.

```{r }
testing_data <- read.csv("pml-testing.csv", na.strings=c("NA",""), strip.white=TRUE)
testing_data <- testing_data[, which(num_of_NA == 0)]
testing_data <- testing_data[, -(1:7)]

predict(model, newdata=testing_data)
```


